---
title: Introduction to writing functions in R 
output: 
  html_document:
    theme: default
    css: custom.css
    includes:
      after_body: footer.html
---

<style>
body{
  font-size: 16px;
  line-height: 24px;
}
.main-container {
  max-width: 1200px;
}
</style>

My personal notes from 
[DataCamp's course](https://www.datacamp.com/courses/introduction-to-writing-functions-in-r)

###### [Back to home](index.html)

### Contents
[Arguments validation](#arguments-validation)  
[Returning multiple values from functions](#returning-multiple-values)

### Arguments validation {#arguments-validation}
The `assertive` package comes with several functions to validate arguments 
within functions.

```{r}
library(assertive)

calc_harmonic_mean <- function(x, na.rm = FALSE) {
  assert_is_numeric(x)
  if(any(is_non_positive(x), na.rm = TRUE)) {
    stop("x contains non-positive values, so the harmonic mean makes no sense.")
  }
  na.rm <- coerce_to(use_first(na.rm), target_class = "logical")
  
  print(1 / mean(1 / x, na.rm = na.rm))
}

calc_harmonic_mean(1:5, na.rm = 1:5)
```

### Returning multiple values from functions {#returning-multiple-values}
Functions can only return one value. If you want to return multiple things, then
you can store them all in a list.

If users want to have the list items as separate variables, they can assign each
list element to its own variable using `zeallot`'s multi-assignment operator, `%<-%`.

Create model object to use in the examples:

```{r}
suppressPackageStartupMessages(library(dplyr))

snake_river_visits <- readRDS('data/snake_river_visits.rds')
model <- lm(n_visits ~ gender + income + travel, snake_river_visits)
```

Returning a list a spliting items via %<-% operator:

```{r}
library(broom)
library(zeallot)

groom_model <- function(model) {
  list(
    model = glance(model),
    coefficients = tidy(model),
    observations = augment(model)
  )
}

# Call groom_model on model, assigning to 3 variables
c(mdl, cff, obs) %<-% groom_model(model)

# See these individual variables
mdl; cff; obs
```

Sometimes you want the return multiple things from a function, but you want the
result to have a particular class (for example, a data frame or a numeric
vector), so returning a list isn't appropriate. This is common when you have a
result plus metadata about the result. (Metadata is "data about the data". For
example, it could be the file a dataset was loaded from, or the username of the
person who created the variable, or the number of iterations for an algorithm to
converge.)

```{r}
pipeable_plot <- function(data, formula) {
  plot(formula, data)
  attr(data, 'formula') <- formula
  invisible(data)
}

plt_dist_vs_speed <- cars %>% 
  pipeable_plot(dist ~ speed)

str(plt_dist_vs_speed)
```

