---
title: Statistics in R
author: My personal knowledge base
output: 
  html_document:
    theme: default
    css: custom.css
    includes:
      after_body: footer.html
---

<style>
body{
  font-size: 16px;
  line-height: 24px;
}
.main-container {
  max-width: 1200px;
}
</style>

###### [Back to home](index.html)

### Contents
- [Models](#models)  
-- [Linear models](#linear-models)  
-- [Non-linear models with GAM](#gam)  
- [Quantiles](#quantiles)

## Models {#models}
### Linear models {#linear-models}
Basic prediction from a linear model:

```{r}
suppressPackageStartupMessages(library(dplyr))

snake_river_visits <- readRDS('data/snake_river_visits.rds')
str(snake_river_visits)
snake_river_explanatory <- readRDS('data/snake_river_explanatory.rds')
str(snake_river_explanatory)

model <- lm(n_visits ~ gender + income + travel, snake_river_visits)

snake_river_explanatory %>%
  mutate(predicted_n_visits = predict(model, .))%>%
  arrange(desc(predicted_n_visits))
```
###### (https://learn.datacamp.com/courses/introduction-to-writing-functions-in-r)

### Non-linear models with GAM {#gam}
When you draw a smooth trend line in plot using `ggplot2::geom_smooth()`, 
you're, in fact, using a generalized additive model (GAM). This sort of model 
is ideal for fitting nonlinear curves. You can use `mgcv::gam()` to run the 
model. Thesyntax for running this GAM takes the following form:

```
gam(response ~ s(explanatory_var1) + explanatory_var2, data = dataset)
```

Here, `s()` means "make the variable smooth", where smooth very roughly means 
nonlinear.

```{r}
library(ggplot2)

df <- readRDS('data/nass.corn.rds')
states_data <- readRDS('data/states.rds')

# get BEA region of each state
df <- df %>%
  inner_join(states_data, by = c('state' = 'name'))

ggplot(df, aes(year, yield_bushels_per_acre)) +
  geom_line(aes(group = state)) +
  geom_smooth() +
  facet_wrap(vars(bea_region))
```

We can see in the plots that the trend lines fit well. So we can use GAM model
and make predictions.

```{r}
suppressPackageStartupMessages(library(mgcv))

bea_regions <- readRDS('data/BEA_regions.rds')$region_name

model <- gam(yield_bushels_per_acre ~ s(year) + bea_region, data = df)

predict_this <- data.frame(
  year = 2050,
  bea_region = bea_regions
)

# Predict the yield
pred_yield <- predict(model, predict_this, type = "response")

predict_this %>%
  mutate(pred_yield = pred_yield)
```
###### (https://learn.datacamp.com/courses/introduction-to-writing-functions-in-r)


## Quantiles {#quantiles}
A custom function ```cut_by_quantile()``` that converts a numeric vector into a categorical variable where quantiles define the cut points:

```{r}
cut_by_quantile <- function(x, n = 5, na.rm = FALSE, labels = NULL, 
                            interval_type = c("(lo, hi]", "[lo, hi)")) {
  interval_type <- match.arg(interval_type)
  probs <- seq(0, 1, length.out = n + 1)
  qtiles <- quantile(x, probs, na.rm = na.rm, names = FALSE)
  right <- switch(interval_type, "(lo, hi]" = TRUE, "[lo, hi)" = FALSE)
  cut(x, qtiles, labels = labels, right = right, include.lowest = TRUE)
}

# Remove the interval_type argument from the call
cut_by_quantile(snake_river_visits$n_visits)
```

